---
title: "Self-Labeling Analysis"
output: html_document
date: "2023-09-26"
---

Pre-registration: https://osf.io/mxnbu/

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## Load packages
library(tidyverse)
library(here)
library(kableExtra)
library(scales)
library(oddsratio)
library(broom)

## Load data
df <- readRDS(here("Clean Data.rds"))

## Set functions
`%+%` <- paste0

percents <- function(x) {
  
  df %>%
    count({{x}}) %>%
    mutate(pct = percent(n / sum(n), .1)) %>%
    kable() %>%
    kable_styling() %>%
    return()
  
}

percents_by_group <- function(x, g) {
  
  df %>%
    count({{g}}, {{x}}) %>%
    mutate(pct = percent(n / sum(n))) %>%
    group_by({{g}}) %>%
    mutate(group_pct = percent(n / sum(n), .1)) %>%
    kable() %>%
    kable_styling() %>%
    return()
  
}
```

# Sample Description

## Self-Labeling

```{r}
percents(depression_self_id)
percents(PHQ8_cutoff)
percents(depression_dx)

percents_by_group(depression_self_id, PHQ8_cutoff)
chisq.test(df$depression_self_id, df$PHQ8_cutoff)

percents_by_group(depression_self_id, depression_dx)
chisq.test(df$depression_self_id, df$depression_dx)
```

## Demographics

```{r}
df %>%
  summarize(mean_age = mean(age),
            sd_age = sd(age)) %>%
  kable() %>%
  kable_styling()

percents(gender)
percents(race_ethnicity)
percents(family_income)
```

# RQ 1: Are demographic characteristics associated with depression self-labeling?

We will assess the relationship between demographic characteristics via logistic regressions where self-labeling is the dependent variable and the demographic variable is the independent variable. We will report odds ratios for each demographic variable on self-labeling both without (4 regressions) and with (4 additional regressions) depression symptom severity as a covariate. 

Missing data: Imputed with missForest algorithm.
Multiple comparison adjustment: False discovery rate with α set to .05

```{r}
demographics <- c(
  "age",
  "gender",
  "race_ethnicity",
  "family_income"
)

rq_1_uncontrolled <- tibble()
rq_1_controlled <- tibble()

for(x in demographics) {
  
  if(x == "race_ethnicity") {
    
    model_data <- df %>%
      filter(!race_ethnicity %in% c("AI/AN non-Hispanic", "NH/PI non-Hispanic"))
    
  } else model_data <- df

  # Set formulas
  formula_uncontrolled <- as.formula("depression_self_id ~ " %+% x)
  formula_controlled <- as.formula("depression_self_id ~ " %+% x %+% "+ PHQ8")
  
  # Run logistic regressions
  glm_uncontrolled <- glm(data = model_data, formula = formula_uncontrolled, family = "binomial")
  glm_controlled <- glm(data = model_data, formula = formula_controlled, family = "binomial")
  
  # Save coefficients and p-values
  coef_uncontrolled <- tidy(glm_uncontrolled)
  coef_controlled <- tidy(glm_controlled)
  
  # Save odds ratios and confidence intervals
  or_uncontrolled <- or_glm(data = model_data, model = glm_uncontrolled, incr = list(age = 1), ci = .95)
  or_controlled <- or_glm(data = model_data, model = glm_controlled, incr = list(age = 1, PHQ8 = 1), ci = .95)
  
  # Combine coefficients, p-values, odds ratios, and confidence intervals
  glm_output_uncontrolled <- inner_join(coef_uncontrolled, or_uncontrolled, by = c("term" = "predictor"))
  glm_output_controlled <- inner_join(coef_controlled, or_controlled, by = c("term" = "predictor"))
  
  # Attach output to larger table
  rq_1_uncontrolled <- rbind(rq_1_uncontrolled, glm_output_uncontrolled)
  rq_1_controlled <- rbind(rq_1_controlled, glm_output_controlled)
  
}

# Adjust p-values
rq_1_uncontrolled$p_adj <- p.adjust(rq_1_uncontrolled$p.value, method = "fdr")
rq_1_controlled$p_adj <- p.adjust(rq_1_controlled$p.value, method = "fdr")

# Print output
kable(rq_1_uncontrolled, caption = "RQ 1, Unadjusted") %>%
  kable_styling()
kable(rq_1_controlled, caption = "RQ 1, Adjusted") %>%
  kable_styling()
```

# RQ 2: What is the relationship between self-labeling as being depressed and self-management strategies (emotion coping strategies, treatment efficacy beliefs)?

We will assess the relationship between self-labeling and self-management strategies using linear regression, where the self-management variable is the dependent variable and depression self-labeling is the independent variable. Again, we will report regression coefficients for self-labeling on each self-management outcome both without (11 regressions) and with (11 additional regressions) depression symptom severity as a covariate.

Missing data: Imputed with missForest algorithm.
Multiple comparison adjustment: False discovery rate with α set to .05

```{r}
# RQ 2
outcomes <- c(
  "perceived_control",
  "CERQ_self_blame",
  "CERQ_other_blame",
  "CERQ_rumination",
  "CERQ_catastrophizing",
  "CERQ_perspective",
  "CERQ_refocusing",
  "CERQ_reappraising",
  "CERQ_acceptance",
  "CERQ_planning",
  "MHSAS_therapy",
  "MHSAS_medication"
)

rq_2_uncontrolled <- tibble()
rq_2_controlled <- tibble()

for(y in outcomes) {
  
  model_data <- df

  # Set formulas
  formula_uncontrolled <- as.formula(y %+% "~ depression_self_id")
  formula_controlled <- as.formula(y %+% "~ depression_self_id + PHQ8")
  
  # Run logistic regressions
  lm_uncontrolled <- lm(data = model_data, formula = formula_uncontrolled)
  lm_controlled <- lm(data = model_data, formula = formula_controlled)
  
  # Save coefficients and p-values
  coef_uncontrolled <- tidy(lm_uncontrolled) %>%
    mutate(ci_low = estimate - (1.96 * std.error),
           ci_high = estimate + (1.96 * std.error)) %>%
    filter(term != "(Intercept)")
  coef_controlled <- tidy(lm_controlled) %>%
    mutate(ci_low = estimate - (1.96 * std.error),
           ci_high = estimate + (1.96 * std.error)) %>%
    filter(term != "(Intercept)")
  
  # Attach DV information
  coef_uncontrolled$dv <- y
  coef_controlled$dv <- y
  
  # Attach output to larger table
  rq_2_uncontrolled <- rbind(rq_2_uncontrolled, coef_uncontrolled)
  rq_2_controlled <- rbind(rq_2_controlled, coef_controlled)
  
}

# Adjust p-values
rq_2_uncontrolled$p_adj <- p.adjust(rq_2_uncontrolled$p.value, method = "fdr")
rq_2_controlled$p_adj <- p.adjust(rq_2_controlled$p.value, method = "fdr")

# Print output
kable(rq_2_uncontrolled, caption = "RQ 2, Unadjusted") %>%
  kable_styling()
kable(rq_2_controlled, caption = "RQ 2, Adjusted") %>%
  kable_styling()
```

# Secondary Analyses

RQ 2: We may explore non-linear relationships between depression self-labeling, symptom severity, and self-management strategies, e.g., by controlling for symptom severity using a quadratic term, or by conducting further analyses that only examine the relationship between self-labeling and self-management strategies among students with a limited range of depression symptom severity (e.g., only students who meet the clinical cutoff for depression; only students  who experience any depression symptoms; only students “at the margin” who are not clearly depressed or not depressed).

```{r}
# The relationship between self-labeling and perceived control appears to be moderated by depression symptom severity
lm(
  data = df,
  formula = perceived_control ~ depression_self_id * PHQ8
) %>%
  summary()
```

# Plots

Self-labeling by age and PHQ-8, smoothed

```{r}
df %>%
  mutate(age_rounded = round(age)) %>%
  filter(PHQ8_cutoff == T,
         age_rounded %in% c(18, 19, 20, 21, 22)) %>%
  group_by(age_rounded) %>%
  summarize(n = n(),
            prop_self_id = mean(depression_self_id),
            sd_self_id = sd(depression_self_id)) %>%
  mutate(se_self_id = sd_self_id / sqrt(n),
         ci_lower = prop_self_id - (1.96 * se_self_id),
         ci_upper = prop_self_id + (1.96 * se_self_id)) %>%
  ggplot(aes(age_rounded, prop_self_id)) +
  geom_point() + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper))
```

Outcomes by self-labeling and PHQ-8, smoothed

```{r}
df %>%
  pivot_longer(cols = c("perceived_control", "CERQ_rumination", "CERQ_catastrophizing", "CERQ_perspective", "CERQ_refocusing", "CERQ_reappraising", "CERQ_planning", "MHSAS_medication")) %>%
  mutate(name = recode(
    name, 
    CERQ_acceptance = "CERQ: Acceptance",
    CERQ_catastrophizing = "CERQ: Catastrophizing",
    CERQ_other_blame = "CERQ: Other-Blame",
    CERQ_self_blame = "CERQ: Self-Blame",
    CERQ_perspective = "CERQ: Putting into Perspective",
    CERQ_planning = "CERQ: Planning",
    CERQ_reappraising = "CERQ: Positive Reappraisal",
    CERQ_refocusing = "CERQ: Positive Refocusing",
    CERQ_rumination = "CERQ: Rumination",
    MHSAS_medication = "Medication Efficacy Beliefs",
    MHSAS_therapy = "Therapy Efficacy Beliefs",
    perceived_control = "Perceived Control"
  )) %>%
  ggplot(aes(x = PHQ8, y = value, group = depression_self_id, color = depression_self_id)) +
  stat_summary(alpha = .25) +
  geom_smooth(method = "loess", formula = "y ~ x", span = 1, se = F) +
  scale_x_continuous(name = "PHQ-8 Sum Score") +
  scale_y_continuous(name = "z-Score") +
  scale_color_discrete(name = "Depression\nSelf-Labeling") +
  facet_wrap(~ name) +
  theme_classic()
```
